---
title: SS201
subtitle: Principles of Economics
author: MAJ Carson Homme
date: AY 23-2

bibliography: references.bib
bibstyle: spphys
# bibstyle options spbasic(default), spphys, spmpsci
output:
  pdf_document:
    citation_package: default
    latex_engine: xelatex
    toc: false
    number_sections: true
  geometry: margin=1in
  fontsize: 11pt
  compact-title: true
header-includes: 
  \usepackage{booktabs}
  \usepackage{tabularx}
  \usepackage{threeparttable}
  \usepackage{bm}
  \DeclareMathOperator{\Var}{Var}
  \usepackage{dcolumn}
  \usepackage{changepage}
  \usepackage{multirow}
  \usepackage[normalem]{ulem}
  \useunder{\uline}{\ul}{}
  \usepackage{fancyhdr}
  \usepackage{graphicx}
  \usepackage{multicol}
  \usepackage[utf8]{inputenc}
  \AtBeginDocument{\let\maketitle\relax}
  \usepackage{enumitem}
  \setlist[enumerate]{nosep}
  \renewcommand{\labelenumi}{\alph{enumi})}
  \usepackage{array}
  \usepackage{tabularray}
  \newcolumntype{C}[1]{>{\centering\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}

  
  # - \usepackage{xr} ## Needs .aux file which knitr automatically removes
  # - \externaldocument[SM-]{../SM/sceptic-SM}
  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	cache = TRUE, 
	echo = FALSE, 
	warning = FALSE,
	message = FALSE,
	fig.align = 'center',
	out.width = '100%',
	dpi=300
	)
```

```{r libs, cache=FALSE, message=FALSE}
library(data.table)
library(ggplot2)
library(forcats)
library(kableExtra)
```


\fancyhf{}
\begin{center}
    \Large
    \textbf{
    \textit{SS201: Principles of Economics} \\ 
    AY 23-2 \\ 
    }
    Lesson 20: Macroeconomic Aggregates
\end{center}
\fancyfoot[C]{\thepage}

# Review {#sec:review}
For each question below, select either True, False, or Uncertain. Then in the space provided, briefly justify your answer (one or two sentences maximum).


\begin{center}
True \hspace{3cm} False \hspace{3cm} Uncertain
\end{center}

1. In the market for wool socks, the number of buyers has increased and technological advancement in production has also occurred. The new equilibrium in the market for wool socks will result in an increased price and quantity.

  \vspace{2cm}

\begin{center}
True \hspace{3cm} False \hspace{3cm} Uncertain
\end{center}

2. In the market for hamburgers, the price elasticity of demand is inelastic at the price point and quantity sold; thus, increasing the price will result in higher revenue.
  
\vspace{2cm}

\begin{center}
True \hspace{3cm} False \hspace{3cm} Uncertain
\end{center}

3. If a tariff is placed on watches, only the price of imported watches will rise by the amount of the tariff and not the price of domestic watches.

\vspace{2cm}

\begin{center}
True \hspace{3cm} False \hspace{3cm} Uncertain
\end{center}

4. If two individuals have different opportunity costs for producing two goods, they can benefit from trade.


\pagebreak

# Bottom Line Up Front {#sec:bluf}
We now begin our study of the larger economy, how it is measured, and some general models that explain economic growth. In assessing differences and how economies are measured, one always must be careful to distinguish between nominal and real values. Economists use measures such as the Gross Domestic Product Deflator and the Consumer Price Index to help identify real changes within the economy. What is worth a dollar today, may or may not be worth a dollar tomorrow.

# Gross Domestic Product {#sec:gdp}
1.	What does Gross Domestic Product (GDP) measure? \vspace{3cm}

2.	What are the two main approaches to calculating GDP? What is their relationship? \vspace{3cm}

3.	What are the components of GDP, using the expenditure method? \vspace{3cm}

4.	What happens to GDP from each of the following activities? What components are affected?
    a. You get a $40 haircut in NYC \vspace{1cm}
    b. You mow your own lawn \vspace{1cm}
    c. You mow your neighbor’s lawn, he mows yours, and you each pay each other $20. \vspace{1cm}
    d. A steel mill sells 10 tons of steel to a Dodge factory for use that year  \vspace{1cm}
    e. You buy $100 worth of Gamestop stock  \vspace{1cm}
    f. Apple produces a laptop that it does not sell  \vspace{1cm}
    g. Apple sells the laptop next year  \vspace{1cm}
    h. You buy a Ferrari \vspace{1cm}
    i. A restaurant buys a new pizza oven \vspace{1cm}
    j. Government pays $1000 in social security to a retiree  \vspace{1cm}
    k. NY state builds a bridge from West Point to Garrison  \vspace{1cm}
 
5.	What are some limitations of GDP as a measure of economic well-being of our country? \vspace{3cm}

6.	What is the difference between Nominal GDP and Real GDP? \vspace{3cm}

\pagebreak

# Price Levels and Inflation {#sec:pl}
1. List and define two ways to measure the price level, and describe how to calculate them. When might each type of measure be appropriate? \vspace{3cm}

Use the below table to answer the following questions. Fiji produces only sugar, vacations, and planes.

\begin{center}
\resizebox{0.95\textwidth}{!}{
                \begin{tabular}{|C{.08\linewidth}|C{.08\linewidth}|C{.08\linewidth}|C{.08\linewidth}|C{.08\linewidth}|C{.08\linewidth}|C{.08\linewidth}|}
\hline
\multicolumn{1}{|p{1.5cm}|}{\centering Year } & \multicolumn{1}{p{1.5cm}|}{\centering Price \\ of  \\ Sugar} & \multicolumn{1}{p{1.5cm}|}{\centering Quantity (tons) \\ of Sugar} & \multicolumn{1}{p{1.5cm}|}{\centering Price \\ of Vacations} & \multicolumn{1}{p{1.5cm}|}{\centering Quantity \\ of Vacations} & \multicolumn{1}{p{1.5cm}|}{\centering Price \\ of a Plane} & \multicolumn{1}{p{1.5cm}|}{\centering Quantity \\ of \\ Planes} \\
\hline
2013	& 5 & 100 & 1200 & 120 & 500,000 & 1  \\
\hline
2014	& 7 & 150 & 1500 & 125 & 550,000 & 1 \\
\hline
2015	& 7 & 150 & 1500 & 125 & 600,000 & 1 \\
\hline
\end{tabular}}
\end{center}

2. Using the above table, fill out the below.
  
  \begin{center}
                \begin{tabular}{|C{.08\linewidth}|C{.08\linewidth}|C{.08\linewidth}|C{.08\linewidth}|}
\hline
\multicolumn{1}{|p{1.5cm}|}{\centering Year} & \multicolumn{1}{p{3.5cm}|}{\centering Nominal \\ GDP} & \multicolumn{1}{p{3.5cm}|}{\centering Real GDP \\ (\$2013)} & \multicolumn{1}{p{3.5cm}|}{\centering GDP \\ Deflator} \\
\hline
2013	& & & \\
\hline
2014	& & & \\
\hline
2015	& & & \\
\hline
\end{tabular}
\end{center}
  
3. Now assume Fiji uses a bundle of 2 tons of sugar and 1 vacation for its CPI index.  What is the market bundle cost for 2013?  For 2014?  For 2015? What is the CPI (2013 base)? Complete the table below.

\begin{center}
                \begin{tabular}{|C{.08\linewidth}|C{.08\linewidth}|C{.08\linewidth}|C{.08\linewidth}|}
\hline
\multicolumn{1}{|p{1.5cm}|}{\centering Year} & \multicolumn{1}{|p{4.5cm}|}{\centering Cost of Basket} & \multicolumn{1}{|p{4.5cm}|}{\centering CPI} \\
\hline
2013	& & \\
\hline
2014	& & \\
\hline
2015	& & \\
\hline
\end{tabular}
\end{center}

\vspace{3cm}
  
4. What is the inflation rate for 2013 to 2014 using the GDP Deflator?  Using the CPI? \vspace{3cm}
  
5. What is the inflation rate for 2014 to 2015 using the GDP Deflator?  Using the CPI? \vspace{3cm}
  
6. 2LT pay in 1980 was \$566.10 and in 2008 it was \$2,469.30. The CPI in 1980 was 41.8 and in 2008 it was 216.6.

    a.	Calculate 1980 and 2008 2LT Pay in 2008 dollars. \vspace{3cm}
    b.	Calculate 1980 and 2008 2LT Pay in 1980 dollars.  \vspace{3cm}
    c.	Which pay is higher in real terms?  \vspace{3cm}
